### 2013-03-26 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
adding spn sparsity plots to spn equations chapter

%%---------------------------------------------------------------------------%%
\section{Spectral Analysis of the $SP_N$ Equations}
\label{sec:spn_spectral_analysis}
Before we can move on to solving the $SP_N$ equations with both
conventional and Monte Carlo methods, we must first understand their
spectral character to ensure the MCSA will apply. As presented in
Chapter~\ref{chap:mc_background}, the spectral radius of the iteration
matrix must be less than unity to ensure convergence and a finite
variance. Therefore, we will compute the spectral radius of the $SP_N$
equations parametrically varying the $SP_N$ order, the $P_N$ order,
the number of energy groups, and upscattering/downscattering between
energy groups.

>>>>

%%---------------------------------------------------------------------------%%
\section{Spectral Analysis of the $SP_N$ Equations}
\label{sec:spn_spectral_analysis}
Before we can move on to solving the $SP_N$ equations with both
conventional and Monte Carlo methods, we must first understand their
spectral character to ensure the MCSA will apply. As presented in
Chapter~\ref{ch:stochastic_methods}, the spectral radius of the
iteration matrix must be less than unity to ensure convergence and a
finite variance. Therefore, we will compute the spectral radius of the
$SP_N$ equations while parametrically varying the $SP_N$ order, the
$P_N$ order, the number of energy groups, and
upscattering/downscattering between energy groups. For each problem in
the parameter study, the values given in
Table~\ref{tab:spn_fixed_parameters} were fixed for each spatial
location and energy group.
\begin{table}[t!]
  \begin{center}
    \begin{tabular}{ll}\hline\hline
      \multicolumn{1}{c}{Parameter}& 
      \multicolumn{1}{c}{Value} \\\hline\hline
      Mesh Element Size & 1.0 \\
      Mesh Elements & $4 \times 4 \times 4$ \\
      Reflecting Boundaries & True \\
      Materials & 1 \\
      Source Strength & 1.0 \\
      Total Cross Section & 5.0 \\
      In-Group Cross Section & 0.25 \\
      Downscatter Cross Section & 1.0 \\
      Upscatter Cross Section & 0.1 \\
      Eigenvalue Solver Tolerance & $1.0\times10^{-8}$ \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Fixed parameters for the $SP_N$ spectral radius
      parameter study.} \textit{The parameters were fixed for each
      spatial location and energy group.}}
  \label{tab:spn_fixed_parameters}
\end{table}
Mesh element sizes were the same in all cardinal directions with
reflecting conditions on each boundary. For the energy parameter, 1
and 10 energy groups were used with upscatter/downscatter varied for
the 10 group case. For the downscatter cases, all groups could scatter
to all lower energy groups with the cross sections given in
Table~\ref{tab:spn_fixed_parameters} while for the upscatter case all
groups could scatter to all other energy groups with different cross
sections for upscatter and downscatter used. A uniform isotropic
source of neutrons is given in each group as well.

The matrices generated by the discretization of this problem are
sparse with a block-based form as dictated by
Eq~(\ref{eq:spn_multigroup_system}). Figure~\ref{fig:group1} gives the
sparsity pattern for a monoenergetic $SP_7$ discretization while
Figures~\ref{fig:group10ds} and \ref{fig:group10us} give the sparsity
patterns for the same problem for 10 energy groups without and with
upscatter respectively. For these figures, the parameters in
Table~\ref{tab:spn_fixed_parameters} with the number of spatial
elements reduced to $2 \times 2 \times 2$ to show the blocks generated
by the discretization.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=3in]{chapters/spn_equations/group1.png}
  \end{center}
  \caption{\textbf{Sparsity pattern for 1-group $SP_7$
      discretization.} \textit{A $2\times 2 \times 2$ element mesh was
      used to show detil of the blocks formed by the discretization.}}
  \label{fig:group1}
\end{figure}
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=3in]{chapters/spn_equations/group10ds.png}
  \end{center}
  \caption{\textbf{Sparsity pattern for 10-group $SP_7$ discretization
      with downscatter only.} \textit{A $2\times 2 \times 2$ element
      mesh was used to show detil of the blocks formed by the
      discretization.}}
  \label{fig:group10ds}
\end{figure}
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=3in]{chapters/spn_equations/group10us.png}
  \end{center}
  \caption{\textbf{Sparsity pattern for 10-group $SP_7$ discretization
      with downscatter and upscatter.} \textit{A $2\times 2 \times 2$ element
      mesh was used to show detil of the blocks formed by the
      discretization.}}
  \label{fig:group10us}
\end{figure}
